A Detailed Analysis of Performance Bottlenecks in MRI Pre-Processing

### 2023-05-05 paper/paper.tex
[ADD] example to run vtune in sigularity + [UPDATE] some sections.

\usepackage{cite}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{graphicx}
\graphicspath{{figures/}}

@@

\subsection{Pipelines}
In this analysis, we aim to cover the pre-processing performance for different MRI modalities. We study the performance bottleneck of large toolboxes and individual pipelines. We use the pipelines from the LivingPark project as a real use case of MRI pre-processing.
% TODO citation for LivingPark project. 
% Q: Ask Tristan about it.

@@

Last, the LivingPark project replicates several papers on Parkinson's Disease with a broad range of analysis. The study uses the following pipelines: FSL SIENA, SPM DARTEL, and SPM DBM.
% TODO Validate which pipelines are used in the project.
% Ideally use all the LivingPark pipelines. If some cannot be used, explain why.

@@

\subsection{Profiling}
We use the Intel VTune profile for our analysis. On the one hand, it has a low performance overhead and offers reporting profiling data at various granularities. On the hand, it requires the debug information of an application to make sense of its profiling data. Otherwise, function names are not available. Therefore, as a first step, we re-compiled each pipeline with debug information in a Docker image. For use on HPC systems, created Singularity images using Docker2Singularity.

Afterwards, we profile a pipeline by mounting the VTune binary to the Singularity image at execution. We store the profiling result for each (pipeline, subject) pair. Combining the Pareto's law and VTune bottom-up report, we identify the top 20\% of function calls that approximately account for 80\% of the pipeline execution time. This heuristic allows us to focus our efforts on the performance critical sections of a pipeline.
% TODO Add code snippet to illustrate VTune profiling of Singularity pipeline.

@@

\subsection{Infrastructure}
% TODO
% Decide between slashbin or compute canada. Most likely going to use CC.

@@

\subsection{Hotspot}
% (1-2 pages)
% TopK (approx. 10) longest makepsan application. The rest can go in supplemental.
% ###    FIGURE    ###
% (Large pipeline)
% fMRIPrep
% sMRIPrep
% DIPY
% 
% (Pipeline components)
% FreeSurfer
% ANTs ??
% SPM DARTEL
% ### END FIGURE ####

% Supplementary figure for the sub-components of large pipelines.

@@

\href{https://github.com/mathdugre/neuro-bottleneck}{https://github.com/mathdugre/neuro-bottleneck}
% TODO Update URL to slashbin org, when created.

>>>>

\usepackage{cite}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{graphicx}
\graphicspath{{figures}}
\usepackage{listings}
\renewcommand{\lstlistingname}{Algorithm}
\usepackage[binary-units=true]{siunitx}

@@

\subsection{Pipelines}
In this analysis, we aim to cover the pre-processing performance for different MRI modalities. We study the performance bottleneck of large toolboxes and individual pipelines. We use the pipelines from the LivingPark project as a real use case of MRI pre-processing.
% TODO citation for LivingPark project. 
% Q: Ask Tristan about it.
% A: GitHub page for now
% -> Added to paperpile

@@

Last, the LivingPark project replicates several papers on Parkinson's Disease with a broad range of analysis. The study uses the following pipelines: FSL SIENA, SPM DARTEL, and SPM pairwise registration.
% TODO Validate which pipelines are used in the project.
% Ideally use all the LivingPark pipelines. If some cannot be used, explain why.

@@

\subsection{Profiling}
We use the Intel VTune profile for our analysis. On the one hand, it has a low performance overhead and offers reporting profiling data at various granularity. On the hand, it requires the debug information of an application to make sense of its profiling data. Otherwise, function names are not available. Therefore, as a first step, we re-compiled each pipeline with debug information in a Docker image. For use on HPC systems, we created Singularity images using Docker2Singularity.

\lstlistingname~\ref{vtune_example} illustrate the profiling of a pipeline by mounting the VTune binary to the Singularity image at execution. We store the profiling result for each (pipeline, subject) pair. Combining the Pareto's law and VTune bottom-up report, we identify the top 20\% of function calls that approximately account for 80\% of the pipeline execution time. This heuristic allows us to focus our efforts on the performance critical sections of a pipeline.

\lstinputlisting[
	language={bash},
	caption={Profiling an application within a Singularity container, using VTune profiler.},
	captionpos=b,
	label=vtune_example,
	numbers=left,
	frame=single,
	basicstyle=\footnotesize,
	numbersep=5pt,
	numberstyle=\tiny\color{gray},
	rulecolor=\color{black},
	tabsize=2,
]{vtune-example.sh}

@@

% TODO explain profiling with single-thread vs. multi-thread.
We set all applications to use a single thread, while profiling. This let

\subsection{Infrastructure}
% TODO
% slashbin description
For our experiments, we used the \textit{Slashbin cluster} at Concordia University. The cluster has eight compute nodes, each configured with two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ \SI{2.10}{\giga\hertz}, \SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs, six \SI{447}{\gibi\byte} SSDs with the XFS file system (no RAID enabled), CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.

Before profiling each application, we transfer the input data from the shared file system to the compute node. After profiling, we transfer back the output to the shared file system. This limits profiling variability due to I/O contingency.

The data generated by the VTune profiler is written to a shared file system, for administrative reasons. Theoretically, this increase the profiling overhead. However, in practice there is no impact since the data generated by the VTune profiler is very small.

@@

\subsection{Hotspot}
% (1-2 pages)
% TopK (approx. 10) longest makepsan application. The rest can go in supplemental.
% ###    FIGURE    ###
% (Large pipeline)
% fMRIPrep
% sMRIPrep
% DIPY
% 
% (Pipeline components)
% SPM DARTEL
% SPM pairwise registration
% ### END FIGURE ####

% Supplementary figure for the sub-components of large pipelines.
% Or include them in #(Pipeline components), if space allows.

@@

\href{https://github.com/mathdugre/mri-bottleneck}{https://github.com/mathdugre/mri-bottleneck}
% TODO Update URL to slashbin org, when created.
